---
pdf_document:
  toc: yes
author: "Andrew D. Johnston"
date: "12/17/2020"
output:
  pdf_document: default
  css: styles_1.css
title: Analysis for the manuscript 'Hepcidin levels can distinguish anemia of chronic
  disease from iron deficiency anemia in a cross-sectional study of hidradenitis suppurativa
  patients'
editor_options:
  chunk_output_type: console
toc: yes
toc_depth: 4
word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache=TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

```{r functions, echo=FALSE, warning=FALSE, message = FALSE}

# Order correlation matrix
reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}
# Get lower triangle of the correlation matrix
get_lower_tri<-function(cormat){
    cormat[upper.tri(cormat)] <- NA
    return(cormat)
}
# Get upper triangle of the correlation matrix
get_upper_tri <- function(cormat){
    cormat[lower.tri(cormat)]<- NA
    return(cormat)
}
```

\newpage

This document details the analysis and generation of figures found in . The dataset related to this article can be found at https://data.mendeley.com/datasets/9x4n84cxxw/1, an open-source online data repository hosted at Mendeley Data Ghias, Mondana; Johnston, Andrew; Babbush, Kayla; Cohen, Steven (2020), “Elevated hepcidin in hidradenitis suppurativa”, Mendeley Data, v1 http://dx.doi.org/10.17632/9x4n84cxxw.1. 

#Overview of Analysis

1. Examine the relationship among acute phase reactants, hepcidin, and disease severity (HS-PGA)
1a. Univariate ordinal regression Hepcidin and HS-PGA

2. Examine the relationship between hepcidin and different anemia classifications.

3. Using Hepcidin to distinguish among IDA, ACD, and ACD/IDA

4. Making Table 1 

5. Making Figure 1

#Curate Data

Load in packages

```{r packages, echo=TRUE, warning=FALSE, message = FALSE}
# load in libraries
library(data.table)
library(ggplot2)
library(ggthemes)
library(reshape2)
library(ggpubr)
library(RColorBrewer)
library(MASS)
library(AER) 
library(scales)
library("flexplot")
library(overlapping)
library(lattice)
library(lubridate)
library(ROCit)
library(pROC)
library(plotROC)
library(OptimalCutpoints)
library(cutpointr)
library(pastecs)
library(Hmisc)
library(mblm)
library(polycor)
library(psych)
library(table1)
library(dunn.test)

# set options
options(scipen=999, stringsAsFactors = FALSE)
```

download the data. 

```{r load-data, echo=TRUE, warning=FALSE, message = FALSE}
hep_dat <- read.csv("Hepcidin Data Reformated_7.31.19.csv", header = T)
tail(hep_dat[1:132,])
hep_dat <- hep_dat[1:131,] # remove blanks rows
hep_dat <- hep_dat[hep_dat$Visit==1,] # use only the 1st visit data
# removing 1 individual without sufficient data, anemic but no ferritin and tsat data
hep_dat <- hep_dat[!(hep_dat$Anemic==1 & (is.na(hep_dat$Tsat) | is.na(hep_dat$Ferritin))),] 
hep_dat$PGA <- as.factor(hep_dat$PGA)
hep_dat$Gender <- as.factor(hep_dat$Gender)
hep_dat_3<- hep_dat
hep_dat_3$Anemic <- as.factor(hep_dat_3$Anemic)
seq_col <- brewer.pal(6,"YlOrRd")
hep_dat_4<-hep_dat_3
hep_dat_4$PGA <- as.numeric(hep_dat_4$PGA)
```


#1 Relationship among hepcidin, acute phase reactants, and disease severity

##1a. Univariate ordinal regression Hepcidin and HS-PGA

Hepcidin levels positively predict HS-PGA in univariate ordinal regression

```{r univar-analyses, echo=TRUE, warning=FALSE, message = FALSE}
ord_uni_PGA_hep <- polr(PGA~Hepcidin, data=hep_dat, Hess = TRUE)
coeftest(ord_uni_PGA_hep) 
summary(ord_uni_PGA_hep)
## store table
(tab_ord_PGA_hep <- coef(summary(ord_uni_PGA_hep)))
## calculate and store p values
p_ord_PGA_hep <- pnorm(abs(tab_ord_PGA_hep[, "t value"]), lower.tail = FALSE) * 2
## combined table
(tab_ord_PGA_hep <- cbind(tab_ord_PGA_hep, "p value" = p_ord_PGA_hep))
## odds ratio
exp(coef(ord_uni_PGA_hep))
## odds ratio with 95% CI
c(exp(coef(ord_uni_PGA_hep)), exp(confint(ord_uni_PGA_hep)))

# Hepcidin and Hurley
table(hep_dat$Hurley)

```

##1b. Correlation among quantitive factors 

Age, BMI, Hb, MCV, PLT,Fe, Hepcidin, Ferritin, CRP, ESR, Tsat, Transferrin are the factors for which we have data points. First, we will assess the distribution of the variables; Only MCV and Transferrin are normally distributed; therefore, will use pearson correlation test (http://www.sthda.com/english/wiki/normality-test-in-r).

```{r normality, echo=TRUE, warning=FALSE, message = FALSE}
colnames(hep_dat)
# Age, BMI, Hb, MCV, PLT,Fe, Hepcidin, Ferritin,CRP, ESR, Tsat, Transferrin
hep_dat_quant <- hep_dat[,c(4,6,8,10:14,16,18,21:22)]
tail(hep_dat_quant)
dim(hep_dat_quant)

l_qqplot <- list()
for (i in 1:ncol(hep_dat_quant)){
  l_qqplot[[i]] <- ggqqplot(hep_dat_quant[,i])
  names(l_qqplot)[i] <- colnames(hep_dat_quant)[i]
}
l_qqplot[[4]] # MCV 
l_qqplot[[12]] # transferrin

shapiro_tests<-NULL
for (i in 1:ncol(hep_dat_quant)){
  shapiro_tests[i] <- unlist(shapiro.test(hep_dat_quant[,i])[2])
  names(shapiro_tests)[i] <- colnames(hep_dat_quant)[i]
}
which(shapiro_tests > 0.05)
```

Will graphically represent the correlation among factors 

```{r heatmap, echo=TRUE, warning=FALSE, message = FALSE}

cormat_quant <- rcorr(x = as.matrix(hep_dat_quant), type="spearman")
cormat_quant <- round(cormat_quant$r,2)

# make corelations between HS-PGA and other variables
cor_hspga <- NULL
j<-1
for (i in c(4,6,8,10:14,16,18,21:22)) {
  temp <- rcorr(hep_dat$PGA, hep_dat[,i], type = "spearman") # 0.7226236
  cor_hspga <- c(cor_hspga,temp$r[1,2])
  names(cor_hspga)[j] <- colnames(hep_dat)[i]
  j<-j+1
}
cor_hspga

cormat_all <- rbind(cormat_quant, cor_hspga)
cormat_all <- cbind(cormat_all, c(cor_hspga,1))

#round
cormat_all <- round(cormat_all,2)

#fix names
rownames(cormat_all)[13] <- "HS-PGA"
colnames(cormat_all)[13] <- "HS-PGA"


# Reorder the correlation matrix
cormat <- reorder_cormat(cormat_all)
# make lower triangle
upper_tri <- get_upper_tri(cormat)

# Melt the correlation matrix
melted_cormat <- melt(upper_tri, na.rm = TRUE)
# Create a ggheatmap
ggheatmap <- ggplot(melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()
# Print the heatmap
print(ggheatmap)

# labeled heatmap
ggheatmap_labeled <- ggheatmap + 
geom_text(aes(Var2, Var1, label = value), color = "black", size = 4) +
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
print(ggheatmap_labeled)

ggheatmap_ordered_notext <- ggheatmap + 
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
print(ggheatmap_ordered_notext)

# supervised clustering to more easily see hepcidin's relationships with other variables
cormat <- reorder_cormat(cormat_all)
rownames(cormat)

# Hepcidin, ferritin, CRP, HS-PGA, ESR, Age, BMI, PLT, Tsat MCV, Hb, Fe, Transferrin
cormat_hep <- cormat[c(6:7,3,2,4,5,12,1,11,8:10,13),c(6:7,3,2,4,5,12,1,11,8:10,13)] 
upper_tri_cormat_hep <- get_upper_tri(cormat_hep)
melted_cormat_hep <- melt(upper_tri_cormat_hep, na.rm = TRUE)


# Create a ggheatmap
ggheatmap_cormat <- ggplot(melted_cormat_hep, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
    name="Pearson\nCorrelation") +
  theme_minimal()+ # minimal theme
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()
# Print the heatmap
print(ggheatmap_cormat)

ggheatmap_supervised_notext <- ggheatmap_cormat + 
theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(),
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+
  guides(fill = guide_colorbar(barwidth = 7, barheight = 1,
                title.position = "top", title.hjust = 0.5))
print(ggheatmap_supervised_notext)

#Correlations with Acute phase reactants:
  # Positive - CRP, ferritin, Hepcidin
  # Negative - Transferrin
```

Ferritin and hepcidin levels are strongly correlated (.73). Hepcidin levels also correlate with markers of inflammation (CRP, r=0.49) and disease-severity (HS-PGA, r=0.45). Of note, hepcidin levels do not reach the r=0.30 cutoff for ESR (r=0.28); perhaps showing that hepcidin is more affected by acute rather chronic inflammation. 

##1c. Correlation among quantitive factors 

Using univariate analysis, do any variables seem to influence hepcidin levels? Checked with non-parametric linear regression using Siegel method (as hepcidin distribution contains outliers) as follows:

```{r hep-univariate, echo=TRUE, warning=FALSE, message = FALSE}

results_hep_mblm <-NULL
j<-1
for(i in c(1:6,8:ncol(hep_dat_quant))){
  col_nam <- colnames(hep_dat_quant)[i]
  hep_no_na <- hep_dat[!is.na(hep_dat[,i]),]
  m2 <- mblm(as.formula(paste0("Hepcidin","~",col_nam)), data=hep_no_na, repeated = TRUE)
  results_hep_mblm <- rbind(results_hep_mblm,summary(m2)$coefficients[2,])
  rownames(results_hep_mblm)[j] <- col_nam
  j<-j+1
}
results_hep_mblm
# Hepcidin not affected by BMI

# Obesity
# how many patients were obese?
mean(!is.na(hep_dat$BMI))
mean(hep_dat$BMI>30, na.rm=TRUE) #65%
wilcox.test(hep_dat$Hepcidin[hep_dat$BMI>=30], hep_dat$Hepcidin[hep_dat$BMI<30])
# No statistically significant difference


# what about non-continuous value, such as sex of individual
biserial(hep_dat$Hepcidin, hep_dat$Gender)
# -0.1007311
# no correlation.

wilcox.test(hep_dat$Hepcidin[hep_dat$Gender==0], hep_dat$Hepcidin[hep_dat$Gender==1])
# statistically significant difference
fig_hep_sex <- ggplot(hep_dat, aes(x=Gender, y=Hepcidin, fill=PGA)) + 
geom_dotplot(binaxis='y', stackdir='center',
               stackratio=1, dotsize=1, col=NA, position = position_jitterd(width = NULL, height = NULL, quad.points = 100, seed = NA)) +
  scale_fill_manual(values=seq_col) +
  theme_minimal() + 
  stat_summary(data = hep_dat_3, mapping = aes(x=as.numeric(Gender), y=Hepcidin), fun.y= "median", geom="point", color="black", inherit.aes = FALSE, shape = 95, size=20) +
  ggtitle("") +
scale_x_discrete( labels = c("0" = "Male","1"="Female"))
fig_hep_sex

#Though the dataset has a majority females.
table(hep_dat$Gender)
```

#2 Relationship between Hepcidin and Anemia 

##2a. Hepcidin levels in anemia vs. non-anemic

```{r hepcidin-anemia-plot, echo=TRUE, warning=FALSE, message = FALSE}
hep_dat_3<- hep_dat
hep_dat_3$Anemic <- as.factor(hep_dat_3$Anemic)

seq_col <- brewer.pal(6,"YlOrRd")

fig_hep_anemia <- ggplot(hep_dat_3, aes(x=Anemic, y=Hepcidin, fill=PGA)) + 
geom_dotplot(binaxis='y', stackdir='center',
               stackratio=1, dotsize=1, col=NA, position = position_jitterd(width = NULL, height = NULL, quad.points = 100, seed = NA)) +
  scale_fill_manual(values=seq_col) +
  theme_minimal() + 
  stat_summary(data = hep_dat_3, mapping = aes(x=as.numeric(Anemic), y=Hepcidin), fun.y= "median", geom="point", color="black", inherit.aes = FALSE, shape = 95, size=20) +
  ggtitle("") +
scale_x_discrete( labels = c("0" = "No","1"="Yes"))
fig_hep_anemia

# is the hepcidin level significantly different 
wilcox.test(hep_dat_3$Hepcidin[hep_dat_3$Anemic=="0"],
            hep_dat_3$Hepcidin[hep_dat_3$Anemic=="1"], conf.int=TRUE)
```

##2b. Hepcidin levels predicting anemia

Likely driven by the high number of ACD individuals (who would have elevated hepcidin)

```{r anemia-hep-regress, echo=TRUE, warning=FALSE, message = FALSE}
binom_anemia_hep <- glm(Anemic ~ Hepcidin, family=binomial(link='logit'), data=hep_dat)

summary(binom_anemia_hep)
anova(binom_anemia_hep, test="Chisq")
# Hepcidin   1   9.6192       112     146.69 0.001926 **
confint(binom_anemia_hep)
exp(coef(binom_anemia_hep)) # exponentiated coefficients
exp(confint(binom_anemia_hep)) # 95% CI for exponentiated coefficients
predict(binom_anemia_hep, type="response") # predicted values
residuals(binom_anemia_hep, type="deviance") # residuals
```

##2c. Prevalence of anemia and iron deficiency 

As previously found in Tennant et al. (1968), there is a high proportion of anemia in our cohort (43%). Tennant et al. (1968) examined 42 patients, finding 10 had "marked" anemia (Hb <10). Additionally, the found ESR to elevated but only had results for 13 patients (11/13 had elevated ESR). Though a large cross-sectional study in Denmark (Miller et al. 2016) demonstrated no difference in Hb level after age-sex-smoking-adjusted analyses between HS patients and 20,780 of the general population. Though it should be noted that they only found 4.65% of their HS patients to have anemia. Another European (Poland) study,  Ponikowska et al 2020 only found 3 HS patients (3/74; 4%) to have anemia. 

Our high prevalence may be present in the Bronx

```{r anemia-prev, echo=TRUE, warning=FALSE, message = FALSE}
table(hep_dat$Anemic)
prop.table(table(hep_dat$Anemic)) #43% of our population
```

A recent (2020) study, Ponikowska et al 2020 "Deranged Iron Status Evidenced by Iron Deficiency Characterizes Patients with Hidradenitis Suppurativa" defined iron deficiency as ferrtin <100ug/L or ferritin 100-299 with TSat <20%. They found that 75% of their population met iron deficient criteria.

```{r Ponikowska-compare, echo=TRUE, warning=FALSE, message = FALSE}
# unfortunately, missing ferritin and Tsat information for 9 non-anemic individuals 
sum(is.na(hep_dat$Ferritin)) #7 
sum(is.na(hep_dat$Tsat)) #9 

hep_dat_compare <- hep_dat[which(!(is.na(hep_dat$Tsat) | is.na(hep_dat$Ferritin))),]
nrow(hep_dat_compare) #104

# Ponikowska study
hep_dat_compare$Ponikowska <- 0
hep_dat_compare$Ponikowska[which(hep_dat_compare$Ferritin<100)] <- 1
hep_dat_compare$Ponikowska[which(hep_dat_compare$Ferritin<299 & hep_dat_compare$Tsat<20)] <- 1

sum(hep_dat_compare$Ponikowska==1) # 89 individuals
prop.table(table(hep_dat_compare$Ponikowska==1))
# 85.57692 % of patient are iron deficient as defined by parameters in the Ponikowska study
```

##2d. Classifying anemia in our cohort

Prior studies classified IDA, IDA/ACD, ACD by the following:

**IDA**
1) absence of inflammation
AND
2) i. T‐sat <20% and ferritin <30 ug/l (van Santen et al, 2011; Scholz et al, 2019)
OR ii. sTfR‐index >=1 mg/ug (van Santen et al, 2011)
OR iii. T‐sat levels <15% together with ferritin <50 ug/l and MCH in the lowest quintile (Scholz et al, 2019 as Thurnnham et al. 2010 showed that mean ferritin concentrations that were 50% (P , 0.001) and 38% (P , 0.002) higher when CRP and AGP were elevated by inflammation, respectively)

**Inflammation** was defined in prior studies as: 
1) RA study, inflammation solely by CRP >10 mg/ml (1mg/L) OR DAS28-ESR (Scholz et al, 2019) 
2) RA study, active inflammation (defined as a CRP level of >= 10 mg/ml or an ESR of >= 30 mm/hour) (van Santen et al, 2011; Khalaf et al.[30996848])
3) IBD study, only using CRP > 5 mg/ml AND clinical disease activity indices, CDAI (Crohn's disease activity index) for CD and MTWAI (Modified Truelove and Witts activity index) for UC (Mecklenburg et al.)

**ACD** 
was defined as the following in van Santen et al, 2011:
1) Presence of inflammation
2) i. transferrin saturation <20% and ferritin level >=100 g/ml
OR ii. sTfR index < 1 mg/micro-g and ferritin level >=30 micro-g/ml

**IDA/ACD**
1) Presence of inflammation
2) i. ferritin < 100 ng/mL and Tsat < 20%
OR ii. 2) sTfR index 1 mg/micro-g.

Notes:
ESR 30 is the upper limit of normal for women.

Our CRP classifiation was as follows:
CRP 0 <1 mg/dL
CRP 1 1-5 mg/dL
CRP 2 >5 mg/dL

Total body iron stores were classified by ferritin class (0 for ferritin<20; 1 for 20=<ferritin<200; 2 for ferritin>=200). 

Our criteria:
IDA 
1) absence of inflammation (CRP <10 mg/mL or ESR <50))
AND
2) i. T‐sat <20% and ferritin <30 ug/l 
OR ii. T‐sat levels <15% together with ferritin <50 ug/l

ACD
1) Presence of inflammation (CRP >=10 mg/mL or ESR >=50)
2) i. transferrin saturation <20% and ferritin level >=100 g/ml
OR ii. Ferritin level >= 200 g/mL

IDA/ACD
1) Presence of inflammation (CRP >=10 mg/mL or ESR >=50)
2) ferritin < 100 ng/mL and Tsat < 20%

```{r anemia-class, echo=TRUE, warning=FALSE, message = FALSE}
# Determining anemic status 
anemia_dat <- hep_dat
anemia_dat$anemia_type <- "Not anemic"

# IDA ferritin <30 ug/l 
anemia_dat$anemia_type[which(anemia_dat$Anemic==1 & ((anemia_dat$Ferritin<30)))] <- "IDA"
nrow(anemia_dat[anemia_dat$anemia_type=="IDA",]) #10

## defining inflammation
#ACD
anemia_dat$anemia_type[which(anemia_dat$Anemic==1 & (anemia_dat$CRP.class>=1 | anemia_dat$ESR>=50) &
                               ((anemia_dat$Tsat<20 & anemia_dat$Ferritin>=100) | anemia_dat$Ferritin>=200 ))] <- "ACD"
nrow(anemia_dat[anemia_dat$anemia_type=="ACD",]) #7 

#ACD/IDA
anemia_dat$anemia_type[which(anemia_dat$Anemic==1 & (anemia_dat$CRP.class>=1 | anemia_dat$ESR>=50) &
                               anemia_dat$Tsat<20 & anemia_dat$Ferritin<100)] <- "ACD/IDA"
nrow(anemia_dat[anemia_dat$anemia_type=="ACD/IDA",]) # 26

# Other anemia 
anemia_dat$anemia_type[which(anemia_dat$anemia_type=="Not anemic" & anemia_dat$Anemic==1)] <- "Other anemia"

table(anemia_dat$anemia_type)
prop.table(table(anemia_dat$anemia_type[anemia_dat$Anemic==1]))

anemia_dat$anemia_type <- factor(anemia_dat$anemia_type, levels = c("IDA", "ACD/IDA", "ACD", "Other anemia", "Not anemic"))
table(anemia_dat$Ferritin.class)
anemia_dat$CRP_class <- anemia_dat$CRP.class
fig_hep_anemia_type <- ggplot(anemia_dat, aes(x=anemia_type, y=Hepcidin, col=PGA, shape=as.factor(Ferritin.class))) + 
  #geom_dotplot(binaxis='y', stackdir='center',
  #           stackratio=1, dotsize=.75, col=NA, position = position_jitterd(width = .2, height = NULL, quad.points = 100,   #seed = NA)) +
  stat_summary(data = anemia_dat, mapping = aes(x=anemia_type, y=Hepcidin), fun.y= "median", geom="point", color="black", inherit.aes = FALSE, shape = 95, size=20) +
  geom_point(position = position_jitterd(width = .3, height = NULL, quad.points = 100,   seed = NA)) +
  scale_color_manual(values=seq_col) +
  theme_minimal() + 
  ggtitle("") +
  scale_y_continuous(trans='log2')
fig_hep_anemia_type
```

#3 Using anemia to distinguish anemia type

##3a Hepcidin distinguishes ACD/IDA and ACD
First, we will examine the ability of hepcidin to discern between ACD/IDA and ACD

```{r roc-plot-1, echo=TRUE, warning=FALSE, message = FALSE}
# ROC of IDA/ACD vs ACD
anemia_dat_roc_1 <- anemia_dat[(anemia_dat$anemia_type=="ACD/IDA" | anemia_dat$anemia_type=="ACD"),]
anemia_dat_roc_1$anemia_type <- factor(anemia_dat_roc_1$anemia_type, levels = c("ACD/IDA", "ACD"))

rocfit_1 <- pROC::roc(anemia_dat_roc_1$anemia_type, anemia_dat_roc_1$Hepcidin)
pROC::auc(rocfit_1)
youden_coord_1 <- pROC::coords(rocfit_1, x = "b")

rocplot_1 <- ggplot(anemia_dat_roc_1, aes(m = Hepcidin, d = as.numeric(anemia_type))) + 
  geom_roc(n.cuts=0)
rocplot_1 <- rocplot_1 + 
  style_roc(theme = theme_minimal, xlab = "1 - Specificity", ylab = "Sensitivity") +
  annotate("text", x = .75, y = .25, 
           label = paste("AUC =", round(calc_auc(rocplot_1)$AUC, 2)))  +
  geom_point(aes(x=1-youden_coord_1$specificity, y=youden_coord_1$sensitivity), colour="red")+
  geom_abline(slope=1, intercept=0, lty="dashed")+
  ggtitle("ACD v ACD/IDA")
rocplot_1

# examining other ROC package visulaizations and computing bootstrapped calculations of the youden index.
roc_1 <- rocit(score = anemia_dat_roc_1$Hepcidin, class = anemia_dat_roc_1$anemia_type,
                       negref = "ACD/IDA") 
plot(roc_1, values=TRUE)

set.seed(200)
ciAUC_boot_1 <- ciAUC(roc_1, level = 0.90, nboot = 1000)
print(ciAUC_boot_1)
youden_coord_1
opt_cut_1 <- cutpointr(anemia_dat_roc_1, Hepcidin, anemia_type, boot_runs = 1000)
plot_cut_boot(opt_cut_1)+theme_minimal()
```

##3a Hepcidin distinguishes IDA and ACD/IDA

Second, we will examine the ability of hepcidin to discern between IDA and ACD/IDA

```{r roc-plot-2, echo=TRUE, warning=FALSE, message = FALSE}
# ROC of IDA/ACD vs IDA
anemia_dat_roc_4 <- anemia_dat[(anemia_dat$anemia_type=="ACD/IDA" | anemia_dat$anemia_type=="IDA"),]
anemia_dat_roc_4$anemia_type <- factor(anemia_dat_roc_4$anemia_type, levels = c("IDA", "ACD/IDA"))

rocfit_4 <- pROC::roc(anemia_dat_roc_4$anemia_type, anemia_dat_roc_4$Hepcidin)
pROC::auc(rocfit_4)
youden_coord_4 <- pROC::coords(rocfit_4, x = "b")

rocplot_4 <- ggplot(anemia_dat_roc_4, aes(m = Hepcidin, d = as.numeric(anemia_type))) + 
  geom_roc(n.cuts=0)
rocplot_4 <- rocplot_4 + 
  style_roc(theme = theme_minimal, xlab = "1 - Specificity", ylab = "Sensitivity") +
  annotate("text", x = .75, y = .25, 
           label = paste("AUC =", round(calc_auc(rocplot_4)$AUC, 2)))  +
  geom_point(aes(x=1-youden_coord_4$specificity, y=youden_coord_4$sensitivity), colour="red")+
  geom_abline(slope=1, intercept=0, lty="dashed")+
  ggtitle("ACD v ACD/IDA")
rocplot_4

# examining other ROC package visulaizations and computing bootstrapped calculations of the youden index.
roc_4 <- rocit(score = anemia_dat_roc_4$Hepcidin, class = anemia_dat_roc_4$anemia_type,
                       negref = "IDA") 
plot(roc_4, values=TRUE)
set.seed(200)
ciAUC_boot_4 <- ciAUC(roc_4, level = 0.90, nboot = 1000)
print(ciAUC_boot_4)
youden_coord_4
opt_cut_4 <- cutpointr(anemia_dat_roc_4, Hepcidin, anemia_type, boot_runs = 1000)
plot_cut_boot(opt_cut_4)+theme_minimal()
```

#4 Making Demographic/blood parameter panel table 

This code chunk is not evaluated but kept for reference for how Table1 was generated.
```{r table-1, echo=TRUE, warning=FALSE, message = FALSE, eval=FALSE}
tab1_dat<-anemia_dat

tab1_dat$Gender <- factor(tab1_dat$Gender, levels = c(0,1), labels = c("Male", "Female"))
label(tab1_dat$Gender) <- "Sex"
units(tab1_dat$Age) <- "Years"
tab1_dat$Race <- factor(tab1_dat$Race, levels=c(1,0,2,3), labels = c("African-American","Caucasian", "Other","Unavailable/Declined"))
label(tab1_dat$Race) <- "Race"
tab1_dat$PGA <- factor(tab1_dat$PGA, levels=c(0,1,2,3,4,5), labels = c("Clear (0)", "Minimal (1)", "Mild (2)", "Moderate (3)", "Severe (4)", "Very Severe (5)"))
label(tab1_dat$PGA) <- "HS-PGA score"
units(tab1_dat$Hepcidin) <- "ng/mL"

units(tab1_dat$Hb) <- "g/dL"
units(tab1_dat$MCV) <- "fL"
units(tab1_dat$PLT) <- "K/uL"
units(tab1_dat$Fe) <- "ug/dL"
units(tab1_dat$Transferrin) <- "mg/dL"
units(tab1_dat$Tsat) <- "%"
units(tab1_dat$Ferritin) <- "ng/mL"
units(tab1_dat$CRP) <- "mg/dL"
units(tab1_dat$ESR) <- "mm/h"

table1(~ Gender + Age + Race + PGA + Hb + MCV + PLT + Fe + Transferrin + Tsat + Ferritin + CRP + ESR | anemia_type, data=tab1_dat, overall="Overall", cont.rmstat=list(c("miss")))

rndr <- function(x, name, ...) {
    if (!is.numeric(x)) return(render.categorical.default(x))
    what <- "Median [Min, Max]"
    parse.abbrev.render.code(c("", what))(x)
}

table1(strata, labels, groupspan=c(1, 3, 1),
       render.continuous=c(.="Mean (CV%)", .="Median [Min, Max]",
                           "Geo. mean (Geo. CV%)"="GMEAN (GCV%)"))

table1(~ Gender + Age + Race + PGA + Hb + MCV + PLT + Fe + Transferrin + Tsat + Ferritin + CRP + ESR | anemia_type, data=tab1_dat, overall="Overall", render = rndr)

table1(~ Gender + Age + BMI+ Race + PGA + Hb + MCV + PLT + Fe + Transferrin + Tsat + Ferritin + CRP + ESR | anemia_type, data=tab1_dat, overall="Overall", render.continuous=c(.="Median [Q1, Q3]"))
```

#5 Figure 1A

This code chunk creates figure 1 Hepcidin levels by anemia type and performs Kruskal-Wallis ANOVA with Dunn’s multiple comparisons test. 
```{r fig-1, echo=TRUE, warning=FALSE, message = FALSE}
anemia_dat_fig1 <- anemia_dat[!(anemia_dat$anemia_type=="Other anemia"),]

# Obtain signficiance 
dunn.test(anemia_dat_fig1$Hepcidin, g=anemia_dat_fig1$anemia_type, method="bonferroni", kw=TRUE, label=TRUE,
          wrap=FALSE, table=TRUE, list=FALSE, rmc=FALSE, alpha=0.05, altp=FALSE)

fig_1 <- ggplot(anemia_dat_fig1, aes(x=anemia_type, y=Hepcidin, col=PGA, shape=as.factor(Ferritin.class))) + 
  #geom_dotplot(binaxis='y', stackdir='center',
  #           stackratio=1, dotsize=.75, col=NA, position = position_jitterd(width = .2, height = NULL, quad.points = 100,   #seed = NA)) +
  stat_summary(data = anemia_dat_fig1, mapping = aes(x=anemia_type, y=Hepcidin), fun.y= "median", geom="point", color="black", inherit.aes = FALSE, shape = 95, size=20) +
  geom_point(size=2.5, position = position_jitterd(width = .4, height = NULL, quad.points = 100,   seed = NA)) +
  scale_color_manual(values=seq_col) +
  ylab("Hepcidin (ng/mL)") +
  xlab("Anemia Type") +
  theme_minimal() + 
  ggtitle("") +
  scale_y_continuous(trans='log2')
fig_1
```

Session Information:
```{r session-Info, echo=TRUE, warning=FALSE, message = FALSE}
sessionInfo()
```
